Detecting the impact of land cover change on observed rainfall.

### 2019-07-05 Manuscripts/PeerJRainfall_Landuse/Rainfall_Landuse.Rmd
changes requested by PeerJ submission figures

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: Sydney Institute of Agriculture, The University of Sydney, NSW 2006
abstract: > 
 Analysis of observational data to pinpoint impact of land cover change on local rainfall is difficult due to multiple environmental factors that cannot be strictly controlled. In this study we use a statistical approach to identify the relationship between removal of tree cover and rainfall with data from best available sources for two large areas in Australia. Gridded rainfall data between 1979 and 2015 was used for the areas, while large scale (exogenous) effects were represented by mean rainfall across a much larger area and climatic indicators, such as Southern Oscillation Index and Indian Ocean Dipole. Both generalised additive modelling and step trend tests were used for the analysis. 
  For a region in south central Queensland, the reported change in tree clearing between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from a bushfire affected region on the border of New South Wales and Victoria suggests significant changes in the rainfall due to changes in tree cover. This indicates the method works better when an abrupt change in the data can be clearly identified. The results from the step trend test also mainly identified a positive relationship between the tree cover and the rainfall at p < 0.1 at the NSW/Victoria region. High rainfall variability and possible regrowth could have impacted the results in the Queensland region.
header-includes: >
 \usepackage{lipsum}
 \usepackage{textcomp, rotating}
 \usepackage[normalem]{ulem}
bibliography: reflist_chap3.bib
csl: peerj.csl
date: "`r Sys.Date()`"
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
  rticles::peerj_article: default
---
```{r packages, echo=F, warnings=F, message=F, eval=T}
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))
suppressWarnings(library(fields))
suppressWarnings(library(graphics))
suppressWarnings(library(maptools))
suppressWarnings(library(plotKML))
suppressWarnings(library(raster))
suppressWarnings(library(reshape))
suppressWarnings(library(rgeos))
suppressWarnings(library(rgdal))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(tidyverse))
suppressWarnings(require(lubridate))
suppressWarnings(library(deseasonalize))
suppressWarnings(library(mgcv))
suppressWarnings(library(ncdf4))
suppressWarnings(require(gridExtra))
suppressWarnings(require(kableExtra))
#library(nnet)
```

```{r setup, warning=F, message=F, echo =F}
# root dir
knitr::opts_knit$set(root.dir =
      # "E:/cloudstor/chun/chun/manuscripts/PeerJRainfall_Landuse")
      "C:/users/rver4657/owncloud/chun/chun/manuscripts/PeerJRainfall_Landuse")
```

@@

include_graphics("figures/map_selreg.png")
```

@@

```{r bushfire, fig.cap = "Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.", out.width="90%", echo=F}
include_graphics("figures/bushfire_nswvic.png")
```

@@

```{r figure3-tc-simple, fig.cap= "The expected evolution of the land surface after trees have been removed in (left) the QLD region and (right) the NSW/VIC region.", out.width="90%", echo=F}
include_graphics("figures/tc_simple.png")
```

@@

## the selected regions for mapping
root_dir <- "e:/cloudstor/chun/chun/data"
aus_reg <- readShapePoly(paste(root_dir,
                               "Geo\\aus_state2011\\STE11aAust.shp",sep="/"),
                         proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

@@

png("figures/bushfire_nswvic.png", width=960, height=960)
  par(mar=c(3,3,1,7),xpd=T)
  plot(e222,border="white", xlab="",ylab="", cex.axis=2)  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=2)
  text(148.4, -36,"Kosciuszko",font=2,cex=2)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=2)
  text(148, -35.5,"NSW",font=2,cex=2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),axis.args=list(cex.axis=1.5),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.07,legend.width=1.0,lwd.poly=0.1, legend.mar=7.1)
legend(148.7,-36.2,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=1.5)
 dev.off()

@@

png("figures/slats.png", width=1200, height = 900)
Data %>% 
  gather(key="Year",value="clearing", `1988-1989`:`2014-2015`) %>%
  filter(Bioregion == "Mulga Lands" |
           Bioregion == "Brigalow Belt") %>%
  ggplot(aes(Year,clearing, fill=Bioregion)) + 
  geom_bar(stat="identity") + ylab("Clearing (ha)") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(size = rel(2),
                                   angle = 45, vjust=0.5),
        axis.text.y = element_text(size = rel(2)),
        axis.title = element_text(size = rel(2), face = "bold"),
        legend.text = element_text(size = rel(2)),
        legend.title = element_text(size = rel(2), face = "bold"))
dev.off()

@@

```{r treeCover,eval=F , echo = F}

rootdir <-"E:/cloudstor/chun2/data/satellite/MOD44/B"

# read in the file list from the directory
x <- dir(path=rootdir,pattern="Percent_Tree_Cover.tif")
# Identify the different states in the images
# NSW
NSW <- x[grep("NSW",x)]
# Qld
Qld <- x[grep("Qld",x)]

img <- list()
# raster and brick
for(i in 1:length(NSW)) {
  img[[i]] <- raster(paste(rootdir,NSW[i],sep="/"))
}

imgQ <- list()
# raster and brick
for(i in 1:length(Qld)) {
  imgQ[[i]] <- raster(paste(rootdir,Qld[i],sep="/"))
}

@@

plotclr <- brewer.pal(8,"PiYG")

png("figures/tc_figs.png", width=1500, height=900)
par(mfrow=c(1,2), mar=c(5,3,5,5))

#par(mar=c(3,9,2,5))
plot(e111, cex.axis=1.5, font.axis=2, ylab="",xlab="")
image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),
      legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",add=T,lwd=3,ylab="")
plot(mdb_reg2,border="grey",lwd=5,add=T,ylab="")
# FULL PERIOD NSW
par(mar=c(5,0,5,11))
plot(e222, cex.axis=1.5, font.axis=2, ylab="",xlab="")
image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),
      legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",add=T, lwd=3,ylab="")
plot(mdb_reg2,border="grey",lwd=5,add=T,ylab="")
image.plot(zlim=c(-4,4), breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.8,
           legend.width=3,
           legend.mar=5.1,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=1,cex=1.5,font=2),
           axis.args=list(cex.axis=1.5),graphics.reset=F)
legend(147.1,-36.2,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)
dev.off()

@@

```{r ts-mean, echo = F, fig.cap = "The deseasonalised and detrended rainfall over the 30 years period in (a) the QLD region and (b) the NSW/VIC region. The vertical red lines indicate the year of 2003, in which the studied land cover changes occurred. A change in the time series data is not obvious before and after the land cover changes.", out.width="90%", echo=F}

include_graphics("figures/Rainfall_resid.png")
```

@@

png("figures/hist_rainfall.png", width=900, height=900)
rain %>%
  ggplot(aes(Rainfall)) + geom_histogram(fill="blue") +
  facet_wrap(~State) + xlab("Monthly Rainfall (mm)") + 
  ylab("Frequency") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

# 
plot_df <- tibble(Dates = rep(qld_mean$Date,2),
                  Resid = c(qld_deseason,nsw_deseason),
                  State = c(rep("Qld",length(qld_deseason)),
                            rep("NSW",length(nsw_deseason))))

png("figures/Rainfall_resid.png", width=960, height=960)
plot_df %>%
  ggplot(aes(Dates,Resid)) + geom_line() + 
  geom_vline(xintercept = as.Date("2003-01-01"), 
             col="red", size=1.5) +
  geom_text(aes(x=as.Date("2004-12-01"),y=150),
            label="2003", col="red", size=7) +
  facet_wrap(~State, ncol = 1) + xlab("Date") +
  ylab("Deseasonalised rainfall for the NSW and Qld locations") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

png("figures/tc_simple.png", width=1200, height=900)
par(mar=c(5,8,4,3), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=2,las=2, cex.lab=1.5)
par(mar=c(5,8,4,3))
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=2, cex.lab=2,las=1.5)
dev.off()

@@

png("figures/hist_sign_LC.png", width=900, height = 900)
par(mfrow=c(2,1))
hist(unlist(lapply(1:length(coef.rain30.nswvic),function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",summary(coef.rain30.nswvic[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="NSW full data period")
hist(unlist(lapply(1:length(coef.rain30.qld4),function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",summary(coef.rain30.qld4[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="Qld full data period")
dev.off()

@@

# Now make the final plot for the paper with both sites and 2009 and all data
  png("figures/Cp_30yrs.png", width=1800, height=900)
  par(mfrow=c(1,2))
  par(mar=c(5,5,3,7))
  # QLD first
  plot(e111,xlab="",ylab="", cex.lab = 2,
       cex.axis=2)
  image(dist_Cp30_qld4,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=5,add=T)
  plot(e111,add=T)
legend(148,-27.5,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)

  # Then NSW
  par(mar=c(5,0,3,9),xpd=T)
  plot(e222,xlab="",ylab="", cex.lab = 2,
       cex.axis=2) #,border="white")
  image(dist_Cp30_nsw,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
  plot(aus_reg2,border="grey75",lwd=2,add=T)
  plot(mdb_reg2,border="grey75",lwd=5,add=T)
  plot(e222,add=T)
  image.plot(zlim=c(0,4),breaks=c(0,1,2,3,4),lab.breaks=c(0,0.05,0.1,0.2,1),legend.only=T,horizontal=F,col=c("red","orange","yellow","light blue"),legend.shrink=0.6,legend.width=1.2,legend.args=list(text="p value",line=1,cex=2,font=2),axis.args=list(cex.axis=1.8))

@@

# show histogram
plot_df <- tibble(`r.squared` = c(rsq_qld,rsq_nsw),
                      state = c(rep("QLD", length(rsq_qld)),
                                rep("NSW/VIC", length(rsq_nsw))))

png("figures/rsq_fullmodel.png", height=900, width=900)
plot_df %>%
  ggplot(aes(`r.squared`*100, fill=state)) + geom_histogram() +
  scale_fill_manual(values = c("blue","red")) +
  xlab(expression(paste("% of variability explained (",r^2,")"))) +
    theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
png("figures/soi_explHistogram.png", width=960, height=960)
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  facet_wrap(~state) +
  xlab(expression(paste("% of variability explained (adj",r^2,")"))) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
 dev.off()

@@

  tr_nswvic.step.new <- raster(nsw_tif)
  tr_nswvic.step.new[] <- nswvic.z.new
plotclt <- rev(brewer.pal(8,"RdYlBu"))

png("figures/step_new.png", width=1800, height=900)
  par(mar=c(5,3,3,7), mfrow=c(1,2), xpd=T)
  plot(e111,xlab="",ylab="", cex.axis=2)#,border="white")
  image(tr_qld.step.new,col=plotclt,
        breaks=c(-3,-1.645,-1.282,-0.842,0,0.842,1.282,1.645,3),legend=F,add=T)
  plot(mdb_reg2,border="grey",add=T,lwd=5)
  plot(e111,add=T)
# image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,0,1,2,3,4),lab.breaks=c(0.00,0.05,0.1,0.2,1,0.2,0.1,0.05,0.00),legend.only=T,horizontal=F,col=plotclt,legend.shrink=0.7,legend.width=0.9,legend.args=list(text=c(expression(paste("level of significance (",alpha,")")),"positive Z'","negative Z'"),side=c(2,3,1),line=0.7,cex=1.5,font=2),axis.args=list(cex.axis=1.5))
legend(147.8,-27.5,c("State","MDB"),col=c("grey","grey"),
       lwd=c(2,5),bty="n",title="Boundary",cex=1.8)

  # NSW
  par(mar=c(5,0,3,10))
  plot(e222,xlab="",ylab="", cex.axis=2)#,border="white")
  image(tr_nswvic.step.new,col=plotclt,breaks=c(-2.3,-1.645,-1.282,-0.842,0,0.842,1.282,1.645,3),legend=F,add=T)
  plot(mdb_reg2,border="grey",add=T,lwd=5)
  plot(aus_reg2,border="grey",add=T, lwd=2)
  plot(e222,add=T)
  par(mar=c(3,3,1,9),xpd=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,0,1,2,3,4),
           lab.breaks=c(0.00,0.05,0.1,0.2,1,0.2,0.1,0.05,0.00),
           legend.only=T,horizontal=F,col=plotclt,
           legend.shrink=0.7,legend.width=1.5,
           legend.args=list(text=c(expression(paste("
                    level of significance (",alpha,")")),"positive Z'",
                           "negative Z'"),
                            side=c(2,3,1),line=0.9,cex=1.5,font=2),
           axis.args=list(cex.axis=1.5))

>>>>

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: School of Life and Environmental Sciences, Sydney Institute of Agriculture, The University of Sydney, New South Wales, Australia.
abstract: > 
 Analysis of observational data to pinpoint impact of land cover change on local rainfall is difficult due to multiple environmental factors that cannot be strictly controlled. In this study we use a statistical approach to identify the relationship between removal of tree cover and rainfall with data from best available sources for two large areas in Australia. Gridded rainfall data between 1979 and 2015 was used for the areas, while large scale (exogenous) effects were represented by mean rainfall across a much larger area and climatic indicators, such as Southern Oscillation Index and Indian Ocean Dipole. Both generalised additive modelling and step trend tests were used for the analysis. 
  For a region in south central Queensland, the reported change in tree clearing between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from a bushfire affected region on the border of New South Wales and Victoria suggests significant changes in the rainfall due to changes in tree cover. This indicates the method works better when an abrupt change in the data can be clearly identified. The results from the step trend test also mainly identified a positive relationship between the tree cover and the rainfall at p < 0.1 at the NSW/Victoria region. High rainfall variability and possible regrowth could have impacted the results in the Queensland region.
header-includes: >
 \usepackage{lipsum}
 \usepackage{textcomp, rotating}
 \usepackage[normalem]{ulem}
bibliography: reflist_chap3.bib
csl: peerj.csl
date: "`r Sys.Date()`"
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article # for using bookdown features like \@ref()
  rticles::peerj_article: default
---
```{r packages, echo=F, warnings=F, message=F, eval=T}
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))
suppressWarnings(library(fields))
suppressWarnings(library(graphics))
suppressWarnings(library(maptools))
suppressWarnings(library(plotKML))
suppressWarnings(library(raster))
suppressWarnings(library(reshape))
suppressWarnings(library(rgeos))
suppressWarnings(library(rgdal))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(tidyverse))
suppressWarnings(require(lubridate))
suppressWarnings(library(deseasonalize))
suppressWarnings(library(mgcv))
suppressWarnings(library(ncdf4))
suppressWarnings(require(gridExtra))
suppressWarnings(require(kableExtra))
#library(nnet)

```{r setup, warning=F, message=F, echo =F}
computer <- "laptop"

# root dir
if (computer == "laptop") {
  knitr::opts_knit$set(root.dir =
      "C:/users/rver4657/owncloud/chun/chun/manuscripts/PeerJRainfall_Landuse")
} else {
  knitr::opts_knit$set(root.dir =
       "E:/cloudstor/chun/chun/manuscripts/PeerJRainfall_Landuse")
}

@@

include_graphics("figures/Fig1.png")
```

@@

```{r bushfire, fig.cap = "Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.", out.width="90%", echo=F}
include_graphics("figures/Fig2.png")
```

@@

```{r figure3-tc-simple, fig.cap= "The expected evolution of the land surface after trees have been removed in (A) the QLD region and (B) the NSW/VIC region.", out.width="70%", echo=F}
include_graphics("figures/Fig3.png")
```

@@

## the selected regions for mapping
if (computer == "laptop") {
  root_dir <- "c:/users/rver4657/owncloud/chun/chun/data"
} else {
  root_dir <- "e:/cloudstor/chun/chun/data"
}
aus_reg <- readShapePoly(paste(root_dir,
                               "Geo\\aus_state2011\\STE11aAust.shp",sep="/"),
                         proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

@@

# bushfire nsw and vic figure
png("figures/Fig2.png", width=1000, height=1000, units="px")
  par(mar=c(3,3,1,7),xpd=T)
  plot(e222,border="white", xlab="",ylab="", cex.axis=2)  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=2)
  text(148.4, -36,"Kosciuszko",font=2,cex=2)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=2)
  text(148, -35.5,"NSW",font=2,cex=2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),axis.args=list(cex.axis=1.5),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.07,legend.width=1.0,lwd.poly=0.1, legend.mar=7.1)
legend(148.7,-36.2,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=1.5)
 dev.off()

@@

# SLATS figure at end of the paper
png("figures/Fig11.png", width=1200, height = 900, units = "px")
Data %>% 
  gather(key="Year",value="clearing", `1988-1989`:`2014-2015`) %>%
  filter(Bioregion == "Mulga Lands" |
           Bioregion == "Brigalow Belt") %>%
  ggplot(aes(Year,clearing, fill=Bioregion)) + 
  geom_bar(stat="identity") + ylab("Clearing (ha)") +
  scale_fill_grey() +
  theme_bw() +
  theme(axis.text.x = element_text(size = rel(2),
                                   angle = 45, vjust=0.5),
        axis.text.y = element_text(size = rel(2)),
        axis.title = element_text(size = rel(2), face = "bold"),
        legend.text = element_text(size = rel(2)),
        legend.title = element_text(size = rel(2), face = "bold"),
  legend.position=c(0.9,0.7))
dev.off()

@@

```{r treeCover,eval=F , echo = F}
if (computer == "laptop") {
  root_dir <- "c:/users/rver4657/owncloud/chun2/data/satellite/MOD44/B"
} else {
  root_dir <- "e:/cloudstor/chun2/data/satellite/MOD44/B"
}

# read in the file list from the directory
x <- dir(path=root_dir,pattern="Percent_Tree_Cover.tif")
# Identify the different states in the images
# NSW
NSW <- x[grep("NSW",x)]
# Qld
Qld <- x[grep("Qld",x)]

img <- list()
# raster and brick
for(i in 1:length(NSW)) {
  img[[i]] <- raster(paste(root_dir,NSW[i],sep="/"))
}

imgQ <- list()
# raster and brick
for(i in 1:length(Qld)) {
  imgQ[[i]] <- raster(paste(root_dir,Qld[i],sep="/"))
}

@@

plotclr <- brewer.pal(8,"PiYG")
# Fig 5 in the paper was tc_figs.png
png("figures/Fig5.png", width=1500, height=900, unit = "px")
par(mfrow=c(1,2), mar=c(5,3,5,5))

#par(mar=c(3,9,2,5))
plot(e111, cex.axis=1.5, font.axis=2, ylab="",xlab="")
image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),
      legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",add=T,lwd=3,ylab="")
plot(mdb_reg2,border="grey",lwd=5,add=T,ylab="")
mtext("A",3, font=2, cex=2.5, line = -5, at=c(146,-25.6))
# FULL PERIOD NSW
par(mar=c(5,0,5,11))
plot(e222, cex.axis=1.5, font.axis=2, ylab="",xlab="")
image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-20,-10,-5,0,5,10,20,70),
      legend=F,add=T)
#plot(e111,add=T)
plot(aus_reg2,border="grey",add=T, lwd=3,ylab="")
plot(mdb_reg2,border="grey",lwd=5,add=T,ylab="")
mtext("B",3, font=2, cex=2.5, line = -5, at=c(147.2,-35.1))
image.plot(zlim=c(-4,4), breaks=c(-4,-3,-2,-1,-0.2,0.2,1,2,3,4),
           lab.breaks=c(-70,-20,-10,-5,"no tree","0 or",5,10,20,70),
           legend.only=T,horizontal=F,
           col=c("#C51B7D","#DE77AE","#F1B6DA","#FDE0EF","white",
                 "#E6F5D0","#B8E186","#7FBC41","#4D9221"),
           legend.shrink=0.8,
           legend.width=3,
           legend.mar=5.1,
           legend.args=list(text=expression(paste(Delta,
                            "tree (% ground)",sep="")),
                            line=1,cex=1.5,font=2),
           axis.args=list(cex.axis=1.5),graphics.reset=F)
legend(147.1,-36.2,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)
dev.off()

@@

```{r ts-mean, echo = F, fig.cap = "The deseasonalised and detrended rainfall over the 30 years period in (a) the QLD region and (b) the NSW/VIC region. The vertical red lines indicate the year of 2003, in which the studied land cover changes occurred. A change in the time series data is not obvious before and after the land cover changes.", out.width="90%"}

include_graphics("figures/Fig4.png")
```

@@

png("figures/hist_rainfall.png", width=900, height=900, 
    units = "px")
rain %>%
  ggplot(aes(Rainfall)) + geom_histogram(fill="blue") +
  facet_wrap(~State) + xlab("Monthly Rainfall (mm)") + 
  ylab("Frequency") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

#

@@

png("figures/Fig3.png", width=1200, height=900)
par(mar=c(5,8,4,3), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n", lwd=2, col = "blue")
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2, lwd=2)
axis(2,at=c(1,2),labels=c("removed","tree"),cex.axis=2,las=2, cex.lab=1.5, lwd=2)
mtext("A",3, font=2, cex=2.5, line = -3, at=1)
par(mar=c(5,8,4,3))
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n", lwd=2, col = "blue")
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2, lwd=2)
axis(2,at=c(1,2),labels=c("removed","tree"),cex.axis=2, cex.lab=2,las=1.5, lwd=2)
mtext("B",3, font=2, cex=2.5, line = -3, at=1)
dev.off()

@@

png("figures/hist_sign_LC.png", width=900, height = 900, units = "px")
par(mfrow=c(2,1))
hist(unlist(lapply(1:length(coef.rain30.nswvic),function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",summary(coef.rain30.nswvic[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="NSW full data period")
hist(unlist(lapply(1:length(coef.rain30.qld4),function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",summary(coef.rain30.qld4[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="Qld full data period")
dev.off()

@@

# Now make the final plot for the paper with both sites and 2009 and all data
  # Figure 8
  png("figures/Fig8.png", width=1800, height=900, units="px")
  par(mfrow=c(1,2))
  par(mar=c(5,5,3,7))
  # QLD first
  plot(e111,xlab="",ylab="", cex.lab = 2,
       cex.axis=2)
  image(dist_Cp30_qld4,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=5,add=T)
  plot(e111,add=T)
  mtext("A",3, font=2, cex=2.5, line = -5, at=c(146,-25.6))
legend(148,-27.5,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)

  # Then NSW
  par(mar=c(5,0,3,9),xpd=T)
  plot(e222,xlab="",ylab="", cex.lab = 2,
       cex.axis=2) #,border="white")
  image(dist_Cp30_nsw,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
  plot(aus_reg2,border="grey75",lwd=2,add=T)
  plot(mdb_reg2,border="grey75",lwd=5,add=T)
  plot(e222,add=T)
  mtext("B",3, font=2, cex=2.5, line = -5, at=c(147.2,-35.7))
  image.plot(zlim=c(0,4),breaks=c(0,1,2,3,4),lab.breaks=c(0,0.05,0.1,0.2,1),legend.only=T,horizontal=F,col=c("red","orange","yellow","light blue"),legend.shrink=0.6,legend.width=1.2,legend.args=list(text="p value",line=1,cex=2,font=2),axis.args=list(cex.axis=1.8))

@@

# show histogram
plot_df <- tibble(`r.squared` = c(rsq_qld,rsq_nsw),
                      state = c(rep("QLD", length(rsq_qld)),
                                rep("NSW/VIC", length(rsq_nsw))))
# Histogram of rsquared of full model
png("figures/Fig6.png", height=1000, width=1000, units = "px")
plot_df %>%
  ggplot(aes(`r.squared`*100, fill=state)) + geom_histogram() +
  scale_fill_manual(values = c("blue","red")) +
  xlab(expression(paste("% of variability explained (",r^2,")"))) +
  ylab("Frequency") +
   theme_bw() +
    theme(axis.text = element_text(size=rel(2.5)),
        strip.text = element_text(size=rel(3)),
        axis.title = element_text(size=rel(3)),
        legend.text = element_text(size=rel(2.5)),
        legend.title = element_text(size=rel(2.5)),
        legend.position = c(0.15, 0.9)) 
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
# SOI explaining histogram
dat_text <- data.frame(
  label = c("A", "B"),
  state   = c("Qld","NSW"),
  x     = c(2.5,2.5),
  y     = c(285,285)
)

png("figures/Fig7.png", width=960, height=960, units="px")
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  xlab(expression(paste("% of variability explained (adj. ",r^2,
                        ")"))) +  
  ylab("frequency") +
  geom_text(data=dat_text,aes(x=x,y=y, label=label), 
            col="black", size=10, fontface="bold") +
  facet_wrap(~state) +
  theme_bw() +
  theme(axis.text = element_text(size=rel(3)),
        strip.text = element_text(size=rel(3)),
        axis.title = element_text(size=rel(3)))
 dev.off()

@@

  tr_nswvic.step.new <- raster(nsw_tif)
  tr_nswvic.step.new[] <- nswvic.z.new
plotclt <- rev(brewer.pal(8,"RdYlBu"))
# new step trend Z scores
png("figures/Fig9.png", width=1800, height=900, units = "px")
  par(mar=c(5,3,3,7), mfrow=c(1,2), xpd=T)
  plot(e111,xlab="",ylab="", cex.axis=2)#,border="white")
  image(tr_qld.step.new,col=plotclt,
        breaks=c(-3,-1.645,-1.282,-0.842,0,0.842,1.282,1.645,3),legend=F,add=T)
  plot(mdb_reg2,border="grey",add=T,lwd=5)
  plot(e111,add=T)
  mtext("A",3, font=2, cex=2.5, line = -5, at=c(146,-25.6))

# image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,0,1,2,3,4),lab.breaks=c(0.00,0.05,0.1,0.2,1,0.2,0.1,0.05,0.00),legend.only=T,horizontal=F,col=plotclt,legend.shrink=0.7,legend.width=0.9,legend.args=list(text=c(expression(paste("level of significance (",alpha,")")),"positive Z'","negative Z'"),side=c(2,3,1),line=0.7,cex=1.5,font=2),axis.args=list(cex.axis=1.5))
legend(147.8,-27.5,c("State","MDB"),col=c("grey","grey"),
       lwd=c(2,5),bty="n",title="Boundary",cex=1.8)

  # NSW
  par(mar=c(5,0,3,10))
  plot(e222,xlab="",ylab="", cex.axis=2)#,border="white")
  image(tr_nswvic.step.new,col=plotclt,breaks=c(-2.3,-1.645,-1.282,-0.842,0,0.842,1.282,1.645,3),legend=F,add=T)
  plot(mdb_reg2,border="grey",add=T,lwd=5)
  plot(aus_reg2,border="grey",add=T, lwd=2)
  plot(e222,add=T)
  mtext("B",3, font=2, cex=2.5, line = -5, at=c(147.2,-35.7))
  par(mar=c(3,3,1,9),xpd=T)
image.plot(zlim=c(-4,4),breaks=c(-4,-3,-2,-1,0,1,2,3,4),
           lab.breaks=c(0.00,0.05,0.1,0.2,1,0.2,0.1,0.05,0.00),
           legend.only=T,horizontal=F,col=plotclt,
           legend.shrink=0.7,legend.width=1.5,
           legend.args=list(text=c(expression(paste("
                    level of significance (",alpha,")")),"positive Z'",
                           "negative Z'"),
                            side=c(2,3,1),line=0.9,cex=1.5,font=2),
           axis.args=list(cex.axis=1.5))
